Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-04-26 WRITEUP/ms.tex
revised sec 7

\subsection{Is the Mass-Size Relation an Observational Artifact?} \label{sec:MR}
\DL{To be rewritten later.}
The MCCs of \flower lie below the locus of $A_V$\eq100\,mag followed
by Milky Way clouds in the mass-size relation shown in \Fig{MR}. This
can be understood by acknowledging the fact that a given MCC in \flower 
does not necessary contain CO\footnote{This is because CO forms deeper in molecular clouds and 
our simulation does not have enough resolution to directly form CO. In fact, to form CO and make
predictions for CO line emission in the work presented by 
\citet{Vallini18a}, we model the internal structure of MCC within each cell in our simulation.}.
That is, the MCCs of \flower lying along the locus of $A_V$\eq4\,mag demonstrates
that the MCCs in our simulation have lower column densities than the
star-forming cores observed in the Milky Way, which are
observationally well-resolved within an MCC, and indeed, have
sufficient column density to form CO.

@@

The apparently large virial parameters seen in the MCCs of \flower can be understood using the Toomre $Q$ parameter (see \Fig{Qeff}). 
We first note that fragmentation {\it can} happen in regions of low $Q$ 
(if we only consider instability against axisymmetric perturbations), 
but further evolution/collapse depends on the dynamics or equation of state of the gas.
%
Such fragmentation is expected to take place at the critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$. 
Second, this fragmentation scale is greater than the typical size of GMCs. 
% 
This could be interpreted as a result of instability setting the scales for fragmentation, 
{\em but the truly star-forming regions correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales}.
%
Thus, the high virial parameter found for most MCCs in \flower 
indicates that some of them are not collapsing structures (but the virial parameter varies depending on the evolutionary stage 
of \flower). Hence, some MCCs are found to lie in regions of $Q_{\rm eff}>1$.
On the other hand, MCCs in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable
and \SF is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, 
where energy quickly dissipates. % with low virial parameters.

@@

%--------------------------------------------------------------------------
%                                Conclusions
%--------------------------------------------------------------------------
\section{Summary}      \label{sec:conclusion}

\DL{Note to self: New observations that are now routinely obtained with ALMA provides 
sensitive and high fidelity images revealing the morphology of gas and dust in \highz galaxies, and possibly on sub-kpc scales. However, 
as shown here, prototypical galaxies at such high redshift with properties similar to \flower\ will have molecular gas structures
on sub-kpc scales, but they are not the truly star-forming clumps. As illustrated by studies in local galaxies, \SF proceed on smaller scales than 
what existing \obs are reaching. While the properties of galaxies at EoR differ from nearby galaxies, we argue that 
the physics behind \SF is the same, and so regions of where \SF truly takes places/unstable to collapse is on even smaller scales. 
This results means that caution should be taken when interpreting observational data, albeit the impressive resolution and sensitivity that are unprecedented. 
Careful consideration of large-scale motions is required when interpreting the observed data in order to understand the origin of star formation activity 
in galaxies at EoR. 
}

\MM{Each item in the summary needs to refer to definite results and
  analysis in the previous paper.  I have noted the items that don't
  seem to me to have that necessary support.}
We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy, \flower,
at the EoR using the state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by
    UV radiation and metal lines, 
and detailed stellar feedback.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities
to identify MCCs and their sub-structures in \flower and its satellites.
We decompose the molecular structures into non-overlapping 
     objects
by identifying a set of 
     density contours with varying cutoff densities at multiple
evolutionary stages.
Such an identification method is 
     the three-dimensional analog
to MCC identification based on 
    intensity levels in surface brightness maps or
    position-position-velocity cubes
of molecular line tracers (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).

We extract properties such as mass, size, Mach number, velocity
dispersion, gas surface density, and virial parameter for each MCC and
we perform a Toomre-$Q$ stability analysis on the galaxy. 
The properties of the MCCs in \flower are compared with observed MCCs in the Milky Way disk, the Galactic center, and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF. This is done for different evolutionary stages of \flower.

We find that the MCCs of \flower 
have highly supersonic
velocity dispersions ($\sigma\simeq$\,200\,\kms) comparable to
those observed in $z$\ssim2 starburst galaxies.
The mass scale of the MCCs is of the order of $10^{5.5-8.5}$\,\Msun,
depending on the chosen H$_{2}$ density cut. 
The $\sim$200\,pc-scale MCCs found with a low density threshold
correspond to the arms of the disk of \flower. These break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
The more massive and bigger MCCs in \flower compared to the Milky Way
likely result from the higher gas mass fraction, surface density, and velocity dispersion,
which set the scale for fragmentation. \MM{This could be quantified in
  the discussion by calculating the Jeans scale for the disk conditions.}
This physical scale is
consistent with what has been found in higher resolution idealized
simulation of isolated galaxies. \MM{Perhaps this could be shown more explicitly?}
Our simulation here confirms these results at EoR and under
the influence of continuous gas accretion from the surrounding IGM (see e.g., \citealt{Klessen10a, Goldbaum11a}).

We compare the dynamics of MCCs in \flower to \obs in the context of
the Larson's relation. \MM{Which one?}
The MCCs of \flower are found to have higher $\sigma$ and $\Sigma$
\MM{compared to Milky Way clouds?}
systematically regardless of the 
density threshold $n_{\rm cut}$ adopted in identifying the structures.
The scaling relations found are insensitive to the various density
cuts.  \MM{This point needs to be made more clearly, probably by
  defining the colors that correspond to the density cuts in the figures.} \DL{They are color-coded by ncut in Fig 5 and 7.}
 The velocity dispersion remains $\gtrsim$100\,\kms even when we increase the density threshold and 
 even for the molecular substructures.  \MM{You never clearly identify
   what you mean by ``molecular substructures.''  Do you just mean
   objects found with high $n_{\rm cut}$?}
 Such a high velocity dispersion likely results from the strong supernova and stellar feedback \flower experienced 
 over the multiple episodes of bursty \SF.   \MM{Or is it from the
   deep potential?  Feedback-driven turbulence famously decays quickly.}

Virial analysis indicates that MCC/arms of the main disk of \flower are unbound, but substructures have
lower virial parameter.
\MM{where is this explicitly discussed?}
 Some of these sub-MCCs have Toomre $Q_{\rm eff}$ parameters consistent with being
gravitationally unstable.  \MM{I don't see this looking at Figure 9.
  Rather the opposite if anything.} This is consistent with the notion that collapsing structures resulting from
gravitational instability occur within globally stable structures, which are
supported by turbulence and rotation on large scale.
We also find $\alpha_{\rm vir}\simeq$\,2 for MCCs in the satellite
galaxies, \MM{I didn't see any explicit discussion of the MCCs in
  satellite galaxies.} which we interpret as
result of the weaker/delayed stellar feedback compared to \flower. 
In particular, some of the MCCs in the satellites are collapsing structures that have not yet formed stars, as confirmed by their
higher gas-to-stellar fractions. \MM{Also not shown so far as I can see.}
In terms of the scaling relations examined, we do not observe any significant
temporal variations in the MCC dynamics over the course of 700\,Myr
traced in our simulation.

Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
Observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial for better understanding
the role of \highz galaxy populations
in the context of galaxy evolution and the ISM physics behind their intense star formation in the early Universe.
Cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and are subject to the sub-grid models adopted,
serve as a useful tool for examining and making predictions on the
morphology and dynamics of 
the molecular ISM of the first galaxies.
%Future higher resolution zoom-in simulations would certainly be useful to understand
%the evolving dynamical properties of the lower-mass molecular structures in
%galaxies at EoR under the influence of gas inflow/outflow (i.e., accounting for energy injection at large scale due to infall/outflow).
These predictions can in turn be
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
enabling us to test our findings and refine future simulations to shed light on the physical processes behind
\SF since the cosmic dark ages.

>>>>

\subsection{Is the Mass-Size Relation an Observational Artifact?} \label{sec:MR}
The MCCs of \flower lie below the locus of $A_V$\eq100\,mag followed
by Milky Way clouds in the mass-size relation shown in \Fig{MR}. This
can be understood by acknowledging the fact that a given MCC in \flower 
does not necessary contain CO\footnote{This is because CO forms deeper in molecular clouds and 
our simulation does not have enough resolution to directly form CO. In fact, to form CO and make
predictions for CO line emission in the work presented by 
\citet{Vallini18a}, we model the internal structure of MCC within each cell in our simulation.}.
That is, the MCCs of \flower lying along the locus of $A_V$\eq4\,mag demonstrates
that the MCCs in our simulation have lower column densities than the
star-forming cores observed in the Milky Way, which are
observationally well-resolved within an MCC, and indeed, have
sufficient column density to form CO.

@@

The large virial parameters seen in some MCCs of \flower can be understood 
by first noting that fragmentation {\em can} happen in regions of low $Q$ 
(if we only consider instability against axisymmetric perturbations), 
but further evolution/collapse depends on the dynamics or equation of state of the gas.
%
Such fragmentation is expected to take place at the critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$. 
Second, this fragmentation scale is greater than the typical size of GMCs. 
% 
This could be interpreted as a result of instability setting the scales for fragmentation, 
{\em but the truly star-forming regions correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales}.
%
Thus, the high virial parameter found for some MCCs in \flower 
indicates that they are not the collapsing structures (but the virial parameter varies depending on the evolutionary stage 
of \flower) and are found in regions of $Q_{\rm eff}>1$.
On the other hand, MCCs in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable
and \SF is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, 
where energy quickly dissipates. % with low virial parameters.

@@

%--------------------------------------------------------------------------
%                                Conclusions
%--------------------------------------------------------------------------
\section{Summary and Conclusions}      \label{sec:conclusion}

Properties of the star-forming ISM of galaxies near the Epoch of Reionization are now within reach with ALMA. 
While it is possible to obtain sensitive and high fidelity imaging that reveals their gas and dust morphology 
on sub-kpc scales and even smaller, such \obs remain time-consuming; logistically, data from multiple cycles pushing to increasing resolution
and sensitivity are needed. As such, observational studies on such scales are still missing in the literature. 
In this work, we aim to understand the origin and dynamical properties of molecular clouds complexes 
in prototypical galaxies at the EoR in numerical simulations 
to provide a framework within which upcoming \obs can be compared against to aid in the interpretation.

We study the dynamics of MCCs in a \z$\sim$\,6 prototypical galaxy, \flower,
and their temporal evolution using the state-of-the-art cosmological zoom-in simulations (\ncode{Serra}),
which include a chemical network to determine the formation of molecular hydrogen, heating and cooling of the ISM by
UV radiation and metal lines, and detailed stellar feedback.
Properties of \flower are briefly summarized in \Sec{sim}.
We use a three-dimensional clump-finding algorithm to identify MCCs and their sub-structures. 
We decompose the molecular structures into non-overlapping objects by identifying a set of 
density contours with varying cutoff densities at multiple evolutionary stages. 
Most notably, the accreting phase, starburst phase, and quiescent phase. 
We extract properties such as mass, size, Mach number, velocity
dispersion, gas surface density, and virial parameter ($M_{\rm cl}, R, \mathcal{M}, 
\sigma, \Sigma_{\rm gas}, \alpha_{\rm vir}$) for each MCC and perform a Toomre-$Q$ stability analysis on \flower.

The sizes of MCCs are dependent on the choice of $n_{\rm cut}$, 
and thus, some of the MCCs within the main disk break into multiple sub-MCCs at higher $n_{\rm cut}$. 
The $\sim$200\,pc-scale MCCs found with a low density threshold
correspond to the arms of the disk of \flower and break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
Overall, the mass of MCCs ranges between $10^{5.5-8.5}$\,\Msun, depending on the chosen density threshold.
This mass range is comparable to those observed in \z\ssim2 galaxies in rest-frame UV and optical wavelengths.
Similarly massive molecular structures (few times 10$^7$\,\Msun) have been 
reported in idealized closed-box isolated galaxy simulations done at higher resolution (e.g., $l_{\rm cell}\simeq3$\,pc in the 
48\,kpc box studied by \citealt{Behrendt16a}). 
However, the IGM and merger and accretion histories are not properly modeled in such simulations. 
That said, the similar mass range found in MCCs of \flower\ is
reassuring --- our results are not far off despite the resolution limit ($l_{\rm cell}\simeq$\,30\,pc).
On the other hand, the gas velocity dispersion $\sigma$ is rather insensitive to the adopted density threshold.

We compare several observable quantities that reflect the dynamics of MCCs in \flower to those
observed in the Milky Way disk, the Galactic center, and 
gas-rich starburst galaxies in the local Universe, and at the peak epoch of cosmic \SF in the context of the Larson's relations. 
This comparison is done for different evolutionary stages of \flower.

MCCs of \flower are found to have velocity dispersion and gas surface density systematically higher than 
Milky Way clouds regardless of the density threshold $n_{\rm cut}$ adopted. 
These MCCs are, in fact, highly supersonic (with high Mach numbers).
Their velocity dispersions are comparable to those observed in $z$\ssim2 starburst galaxies ($\sigma\simeq200$\,\kms); however, 
comparing the velocity dispersions of MCCs from bulk motions versus turbulence 
indicates that large-scale rotation is an important contributor, even in the 
accreting phase of \flower, which displays a highly disturbed morphology. 
This poses a challenge to the interpretation of spatially resolved molecular gas structures observed in \highz galaxies to-date. 
In particular, any velocity dispersions measured on similar scales as here (even down to $\sim$\,40\,pc) 
should {\em not} be taken as a direct evidence that \highz molecular gas structures are intrinsically more turbulent than those in the local ISM, 
unless contributions from bulk motions are properly accounted for.
%This means that a lot of the velocity dispersion infer on even the few-tens of parsec scales have a non-negligible contribution 
%from rotation (cf. solely random motion). Therefore it is not fair to say that \highz molecular gas ``clumps`` (or whatever name used 
%for structures in the literature) are more turbulent based on solely measured velocity dispersion.

We do not observe any significant temporal variations in the MCC dynamics over the course of 700\,Myr
traced in our simulation, except in the virial parameter, which is the lowest in the pre-starburst phase. 
The virial parameter increases after a starburst as gas is self-regulated by feedback.

% Toomre:
Close resemblance of $Q_{\rm eff}$ and $Q_{\rm star}$ maps 
indicates that contributions from the stellar component play an important role in governing the stability of the MCCs against
axisymmetric perturbations, especially in the central part of \flower. Similarly, stabilizing effect due to the thickness of its disk is also 
non-negligible. Our Toomre-Q analysis thus illustrates the importance of accounting for stellar contribution 
and disk thickness when examining the stability of 
molecular gas structures, especially in relatively evolved and enriched systems at high redshift that are preferentially being observed now.

We perform virial analysis, as motivated by \obs, where the virial parameter is used as a surrogate to comparing the potential to kinetic energies of 
molecule gas structures. The high virial parameter found for some MCCs in \flower 
indicates that they are not the collapsing structures and are found in regions of $Q_{\rm eff}>1$.
On the other hand, MCCs in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable
and \SF is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, 
where energy quickly dissipates. % with low virial parameters.
This is consistent with the notion that collapsing structures resulting from
gravitational instability occur within globally stable structures, which are
supported by turbulence and rotation on large scale.
%% 
%Our results suggest that the more massive and bigger MCCs compared to the Milky Way
%manifest from the higher gas mass fraction, surface density, and velocity dispersion, but \SF takes place on much smaller scales. 
%Low-$Q$ sets the regions that fragments 
%and within those regions where alphavir is low enough is where structures to further collapse to form stars. 
This also implies \obs on spatially resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores),
and thus, \SF in the first galaxies.

% end
Determining the multi-phase ISM properties of early galaxies
is a critical piece to understanding the evolution and
assembly history of galaxies, since they set the pace
for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
While observations leveraging the combination of spatial-spectral imaging of
multi-band continuum and spectral line emission are crucial, 
we shown in this work that prototypical galaxies at the EoR with properties similar to \flower
are expected to have molecular gas structures on sub-kpc scales; however, these are not 
the truly star-forming regions. The linewidth of molecular gas structures probed on these scales partially results from 
large-scale motions even in the dispersion dominated stage. 
This results means that caution should be taken when interpreting observational data, albeit the impressive resolution and sensitivity that are unprecedented. 
Careful consideration of large-scale motions is required when interpreting the observed data in order to understand the origin of star formation activity 
in galaxies at EoR. 
It is thus currently unclear whether \highz molecular gas structures are intrinsically more turbulent than those in the local ISM. 
This will become clearer in the next few years as deep and high resolution imaging of molecular gas at \highz are obtained with ALMA and NOEMA.

On the other hand, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy
statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining 
and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.
These predictions can in turn be 
tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
enabling us to test our findings and refine future simulations to shed light on the physical processes behind
\SF since the cosmic dark ages.
